---
title: "dominicks cereal vizualisation"
author: "Jei Pratheesh"
date: "2025-04-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### loading required libraries

```{r}
library(tidyverse)
library(ggthemes)
library(scales)
```

### Loading the dataset

```{r}
# Load the cleaned dataset
cereals <- read_csv("cleaned_cereal_data.csv")
```

# Graphic 1 – Elementary: Units Sold by Promotion Type

```{r}
ggplot(cereals, aes(x = SALE)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Units Sold by Promotion Type",
       x = "Promotion Type",
       y = "Number of Records (Proxies Unit Sales)") +
  theme_minimal()
```

# Graphic 2 – Intermediate: Weekly Sales Trend of a Top UPC

```{r}
top_upc <- cereals %>%
  group_by(UPC) %>%
  summarise(total_move = sum(MOVE)) %>%
  slice_max(order_by = total_move, n = 1) %>%
  pull(UPC)

cereals %>%
  filter(UPC == top_upc) %>%
  group_by(WEEK) %>%
  summarise(weekly_units = sum(MOVE)) %>%
  ggplot(aes(x = WEEK, y = weekly_units)) +
  geom_line(color = "tomato", size = 1) +
  labs(title = paste("Sales Spike Patterns Over Time for Best-Selling Product UPC:", top_upc),
       x = "Week",
       y = "Units Sold") +
  theme_minimal()
```

# Graphic 3 – Intermediate: Price Distribution by Promotion Type

```{r}
ggplot(cereals, aes(x = SALE, y = PRICE)) +
  geom_boxplot(fill = "skyblue") +
  labs(title = "Price Distribution by Promotion Type",
       x = "Promotion Type",
       y = "Price") +
  theme_minimal()
```

# Graphic 4 – Overall: Price vs. Profit by Promotion

```{r}
ggplot(cereals, aes(x = PRICE, y = PROFIT, color = SALE)) +
  geom_point(alpha = 0.6) +
  labs(title = "Price vs. Profit Colored by Promotion",
       x = "Price",
       y = "Profit") +
  theme_minimal()
```

# Graphic 5 – Multi-Component: Weekly Sales for Top 4 UPCs

```{r}
top_upcs <- cereals %>%
  group_by(UPC) %>%
  summarise(total_move = sum(MOVE)) %>%
  slice_max(order_by = total_move, n = 4) %>%
  pull(UPC)

cereals %>%
  filter(UPC %in% top_upcs) %>%
  group_by(UPC, WEEK) %>%
  summarise(weekly_units = sum(MOVE)) %>%
  ggplot(aes(x = WEEK, y = weekly_units)) +
  geom_line() +
  facet_wrap(~ UPC) +
  labs(title = "Weekly Sales for Top 4 UPCs",
       x = "Week",
       y = "Units Sold") +
  theme_minimal()
```

# Graphic 6 – Multi-Component: Heatmap of Average Profit by Store and Promotion

```{r}
cereals %>%
  group_by(STORE, SALE) %>%
  summarise(avg_profit = mean(PROFIT, na.rm = TRUE)) %>%
  ggplot(aes(x = SALE, y = factor(STORE), fill = avg_profit)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(title = "Average Profit by Store and Promotion Type",
       x = "Promotion Type",
       y = "Store") +
  theme_minimal()
```

# Graphic 1: Top 10 UPCs by Total Profit (Intermediate-Level)

```{r}
top_upcs <- cereals %>%
  group_by(UPC) %>%
  summarise(Total_Profit = sum(PROFIT, na.rm = TRUE)) %>%
  arrange(desc(Total_Profit)) %>%
  slice_max(order_by = Total_Profit, n = 10)

ggplot(top_upcs, aes(x = reorder(as.factor(UPC), Total_Profit), y = Total_Profit)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  scale_y_continuous(labels = scales::comma) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Top 10 UPCs by Total Profit",
       x = "UPC",
       y = "Total Profit ($)") +
  theme_minimal()
```

# Graphic 2: Promotion Effectiveness by Average Sales (Intermediate-Level)

```{r}
# Map promotion codes to full names
promo_labels <- c(B = "Bonus Buy", C = "Coupon", G = "Grocery Ad", L = "Local Ad", S = "Simple price reduction")
cereals$PROMO_FULL <- recode(cereals$PROMO_TYPE, !!!promo_labels)

cereals %>%
  filter(PROMO_FULL != "None") %>%
  group_by(PROMO_FULL) %>%
  summarise(Avg_Sales = mean(CALC_SALES, na.rm = TRUE)) %>%
  ggplot(aes(x = PROMO_FULL, y = Avg_Sales, fill = PROMO_FULL)) +
  geom_col() +
  labs(title = "Average Sales by Promotion Type",
       x = "Promotion Type",
       y = "Average Calculated Sales ($)") +
  theme_minimal()
```

# Graphic 3: Simplified Price vs Profit (Elementary-Level)

```{r}
# Group PRICE into High, Medium, Low buckets instead of bins
max_price <- max(cereals$PRICE, na.rm = TRUE)
cereals <- cereals %>% mutate(
  PRICE_GROUP = case_when(
    PRICE < 2 ~ "Low",
    PRICE >= 2 & PRICE < 5 ~ "Medium",
    PRICE >= 5 & PRICE <= ceiling(max_price) ~ "High",
    TRUE ~ NA_character_
  )
)

# Summarise by these price groups
price_group_summary <- cereals %>%
  filter(!is.na(PRICE_GROUP)) %>%
  group_by(PRICE_GROUP) %>%
  summarise(Average_Profit = mean(PROFIT, na.rm = TRUE))

ggplot(price_group_summary, aes(x = PRICE_GROUP, y = Average_Profit)) +
  geom_col(fill = "darkgreen") +
  labs(title = "Average Profit by Price Category",
       x = "Price Category",
       y = "Average Profit ($)") +
  theme_minimal()
```

# Graphic 4: Sales Spike Rates by Promotion (Intermediate-Level)

```{r}
cereals$PROMO_FULL <- recode(cereals$PROMO_TYPE, !!!promo_labels)

cereals %>%
  filter(PROMO_FULL != "None") %>%
  group_by(PROMO_FULL) %>%
  summarise(Spike_Rate = mean(SALES_SPIKE, na.rm = TRUE)) %>%
  ggplot(aes(x = PROMO_FULL, y = Spike_Rate, fill = PROMO_FULL)) +
  geom_col() +
  labs(title = "Sales Spike Probability by Promotion Type",
       x = "Promotion Type",
       y = "Spike Rate (Top 10% of Sales)") +
  theme_minimal()
```

# Graphic 5: Weekly Profit Trend – Store 51 (Overall-Level)

```{r}
# Automatically find the store with highest total profit
best_store <- cereals %>%
  group_by(STORE) %>%
  summarise(Total_Profit = sum(PROFIT, na.rm = TRUE)) %>%
  arrange(desc(Total_Profit)) %>%
  slice(1) %>%
  pull(STORE)

cereals %>%
  filter(STORE == best_store) %>%
  group_by(WEEK) %>%
  summarise(Weekly_Profit = sum(PROFIT, na.rm = TRUE)) %>%
  ggplot(aes(x = WEEK, y = Weekly_Profit)) +
  geom_line(color = "darkblue") +
  labs(title = paste("Weekly Profit Trend – Best Store -", best_store),
       x = "Week",
       y = "Profit ($)") +
  theme_minimal()
```

# Graphic 6: Profit by Week Group and Promotion Type (Multi-Component, Overall)

```{r}
cereals$PROMO_FULL <- recode(cereals$PROMO_TYPE, !!!promo_labels)

cereals %>%
  filter(PROMO_FULL != "None") %>%
  group_by(WEEK_GROUP, PROMO_FULL) %>%
  summarise(Avg_Profit = mean(PROFIT, na.rm = TRUE)) %>%
  ggplot(aes(x = WEEK_GROUP, y = Avg_Profit, fill = PROMO_FULL)) +
  geom_col(position = "dodge") +
  labs(title = "Profit by Week Group and Promotion Type",
       x = "Time Period",
       y = "Average Profit ($)") +
  theme_minimal()
```

