---
title: "INFS 5116 - Data Visualization Project"
output: PDF
Author: Jei Pratheesh
---

### ===============================
### loading required libraries
### ===============================

```{r}
library(tidyverse)
library(scales)
```

### ===============================
### Loading the dataset
### ===============================

```{r}
# Load the cleaned dataset
cereals <- read_csv("cleaned_cereal_data.csv")
```

# --------------------------------------
# Graphic 1 – Elementary: Units Sold by Promotion Type
# --------------------------------------

```{r}
ggplot(cereals, aes(x = SALE)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Units Sold by Promotion Type",
       x = "Promotion Type",
       y = "Number of Records (Proxies Unit Sales)") +
  theme_minimal()
```

# --------------------------------------
# Graphic 2 – Intermediate: Weekly Sales Trend of a Top UPC
# --------------------------------------

```{r}
top_upc <- cereals %>%
  group_by(UPC) %>%
  summarise(total_move = sum(MOVE)) %>%
  slice_max(order_by = total_move, n = 1) %>%
  pull(UPC)

cereals %>%
  filter(UPC == top_upc) %>%
  group_by(WEEK) %>%
  summarise(weekly_units = sum(MOVE)) %>%
  ggplot(aes(x = WEEK, y = weekly_units)) +
  geom_line(color = "tomato", size = 1) +
  labs(title = paste("Sales Spike Patterns Over Time for Best-Selling Product UPC:", top_upc),
       x = "Week",
       y = "Units Sold") +
  theme_minimal()
```


# --------------------------------------
# Graphic 3 – Intermediate: Price Distribution by Promotion Type
# --------------------------------------

```{r}
ggplot(cereals, aes(x = SALE, y = PRICE)) +
  geom_boxplot(fill = "skyblue") +
  labs(title = "Price Distribution by Promotion Type",
       x = "Promotion Type",
       y = "Price") +
  theme_minimal()
```

# --------------------------------------
# Graphic 4 – Overall: Price vs. Profit by Promotion
# --------------------------------------

```{r}
ggplot(cereals, aes(x = PRICE, y = PROFIT, color = SALE)) +
  geom_point(alpha = 0.6) +
  labs(title = "Price vs. Profit Colored by Promotion",
       x = "Price",
       y = "Profit") +
  theme_minimal()
```

# --------------------------------------
# Graphic 5 – Multi-Component: Weekly Sales for Top 4 UPCs
# --------------------------------------

```{r}
top_upcs <- cereals %>%
  group_by(UPC) %>%
  summarise(total_move = sum(MOVE)) %>%
  slice_max(order_by = total_move, n = 4) %>%
  pull(UPC)

cereals %>%
  filter(UPC %in% top_upcs) %>%
  group_by(UPC, WEEK) %>%
  summarise(weekly_units = sum(MOVE)) %>%
  ggplot(aes(x = WEEK, y = weekly_units)) +
  geom_line() +
  facet_wrap(~ UPC) +
  labs(title = "Weekly Sales for Top 4 UPCs",
       x = "Week",
       y = "Units Sold") +
  theme_minimal()
```

# --------------------------------------
# Graphic 6 – Multi-Component: Heatmap of Average Profit by Store and Promotion
# --------------------------------------

```{r}
cereals %>%
  group_by(STORE, SALE) %>%
  summarise(avg_profit = mean(PROFIT, na.rm = TRUE)) %>%
  ggplot(aes(x = SALE, y = factor(STORE), fill = avg_profit)) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(title = "Average Profit by Store and Promotion Type",
       x = "Promotion Type",
       y = "Store") +
  theme_minimal()
```